## Inference for a Proportion

### Inference for a Proportion: Frequentist Approach

**EXAMPLE: RU-486 is claimed to be an effective "morning after" contraceptive pill, but is it really effective? ** 

**DATA: A total of 40 women came to a health clinic asking for emergency contraception (usually to prevent pregnancy after unprotected sex). They were randomly assigned to RU-486 (treatment) or standard therapy (control), 20 in each group. In the treatment group, 4 out of 20 became pregnant. In the control group, the pregnancy rate is 16 out of 20.**

**QUESTION: How strongly do these data indicate that the treatment is more effective than the control?**

To simplify the framework, let's make it a one proportion problem and just consider the 20 total pregnancies because the two groups have the same sample size. If the treatment and control are equally effective, then the probability that a pregnancy comes from the treatment group ($p$) should be 0.5. If RU-486 is more effective, then the probability that a pregnancy comes from the treatment group ($p$) should be less than 0.5.

Therefore, we can form the hypotheses as below:

* $p =$ probability that a given pregnancy comes from the treatment group

* $H_0: p = 0.5$ (no difference, a pregnancy is equally likely to come from the treatment or control group)

* $H_A: p < 0.5$ (treatment is more effective, a pregnancy is less likely to come from the treatment group)
 
A p-value is needed to make an inference decision with the frequentist approach. The definition of p-value is the probability of observing something *at least* as extreme as the data, given that the null hypothesis ($H_0$) is true. "More extreme" means in the direction of the alternative hypothesis ($H_A$).

Since $H_0$ states that the probability of success (pregnancy) is 0.5, we can calculate the p-value from 20 independent Bernoulli trials where the probability of success is 0.5. The outcome of this experiment is 4 successes in 20 trials, so the goal is to obtain 4 or fewer successes in the 20 Bernoulli trials. 

This probability can be calculated exactly from a binomial distribution with $n=20$ trials and success probability $p=0.5$. Assume $k$ is the actual number of successes observed, the p-value is 

$$P(k \leq 4) = P(k = 0) + P(k = 1) + P(k = 2) + P(k = 3) + P(k = 4)$$.

```{r Bernoulli20, echo=TRUE}
sum(dbinom(0:4, size = 20, p = 0.5))
```

According to $\mathsf{R}$, the probability of getting 4 or fewer successes in 20 trials is 0.0059. Therefore, given that pregnancy is equally likely in the two groups, we get the chance of observing 4 or fewer preganancy in the treatment group is 0.0059. With such a small probability, we reject the null hypothesis and conclude that the data provide convincing evidence for the treatment being more effective than the control. 

### Inference for a Proportion: Bayesian Approach

This section uses the same example, but this time we make the inference for the proportion from a Bayesian approach. Recall that we still consider only the 20 total pregnancies, 4 of which come from the treatment group. The question we would like to answer is that how likely is for 4 pregnancies to occur in the treatment group. Also remember that if the treatment and control are equally effective, and the sample sizes for the two groups are the same, then the probability ($p$) that the pregnancy comes from the treatment group is 0.5. 

Within the Bayesian framework, we need to make some assumptions on the models which generated the data. First, $p$ is a probability, so it can take on any value between 0 and 1. However, let's simplify by using discrete cases -- assume $p$, the chance of a pregnancy comes from the treatment group, can take on nine values, from 10%, 20%, 30%, up to 90%. For example, $p = 20%$ means that among 10 pregnancies, it is expected that 2 of them will occur in the treatment group. Note that we consider all nine models, compared with the frequentist paradigm that whe consider only one model. 

Table \@ref(tab:RU-486prior) specifies the prior probabilities that we want to assign to our assumption. There is no unique correct prior, but any prior probability should reflect our beliefs prior to the experiement. The prior probabilities should incorporate the information from all relevant research before we perform the current experiement.

```{r RU-486prior, echo = FALSE}

p <- seq(0.1, 0.9, 0.1)
prior <- c(rep(0.06, 4), 0.52, rep(0.06, 4))
likelihood <- dbinom(4, size = 20, prob = p)

# posterior
numerator <- prior * likelihood
denominator <- sum(numerator)
posterior <- numerator / denominator
# sum(posterior)

temp <- matrix(c(p,
                 prior,
                 likelihood), 
               nrow = 3, byrow = TRUE)
rownames(temp) <- c("Model ($p$)", "Prior $P(model)$", 
                    "Likelihood $P(data | model)$")

knitr::kable(
  x = temp,
  booktabs = TRUE,
  caption = 'Prior, likelihood, and posterior probabilities for each of the 9 models',
  digits = 4
)
```

This prior incorporates two beliefs: the probability of $p = 0.5$ is highest, and the benefit of the treatment is symmetric. The second belief means that the treatment is equally likely to be better or worse than the standard treatment. Now it is natural to ask how I came up with this prior, and the specification will be discussed in detail later in the course.

Next, let's calculate the likelihood -- the probability of observed data for each model considered. In mathematical terms, we have

$$ P(\text{data } | \text{ model}) = P(k = 4 | n = 20, p)$$

The likelihood can be computed as a binomial with 4 successes and 20 trials with $p$ is equal to the assumed value in each model. The values are listed in Table \@ref(tab:RU-486prior).

```{r progress, echo=FALSE}
# Now on slide #6, video 2:03.
# Lots of problems with the table settings!
```


### Effect of Sample Size on the Posterior


