## Inference and Decision-Making with Multiple Parameters

This section is focused on the extending the Normal-Normal conjugate family introduced in \@ref(sec:normal-normal) to the problem of inference in a Normal population with an unknown mean and variance.  We will introduce the Normal-Gamma conjugate family for inference about the unknown mean and variance and will present Monte Carlo simulation for inference about functions of the parameters as well as sampling from predictive distributions, which can assist with prior elucidation. For situations when limited prior information is available, we discuss a limiting case of the Normal-Gamma conjugate family, leading to priors that can be used for a reference analysis.  Finally, we will show how to create a more flexible and robust prior distribution by using  mixtures of the  Normal-Gamma conjugate prior.  For inference in this case we will introduce Markov Chain Monte Carlo, a powerful simulation method for Bayesian inference.   

It is assumed that the readers have mastered the concepts of one-parameter Normal-Normal conjugate priors. Calculus is not required for this section; however, for those who are comfortable with calculus and would like to go deeper, we shall present starred sections with more details on the derivations. 

### The Normal-Gamma Conjugate Family {#sec:normal-gamma}


In \@ref(sec:normal-normal) we described the normal-normal conjugate  family for inference about an unknown mean $\mu$ with a known standard deviation $\sigma$ when the data were assumed to be a random sample from a normal population. In this section we will introduce the normal-gamma conjugate family for the common situation when $\sigma$ is unknown.  As both $\mu$ and $\sigma^2$ unknown, we will need to specify a **joint** prior distribution to describe our prior uncertainty about them.

**Sampling Model**

Recall that a conjugate pair is a sampling model for the data and prior distribution for the unknown parameters such that the posterior distribution is in the same family of distributions as the prior distribution. 
We will assume that the data are a random sample of size $n$ from a normal population with mean $\mu$ and variance $\sigma^2$; the following is a mathematical shorthand to represent this distribution assumption 

$$\begin{aligned}
Y_1, \ldots Y_n  \iid
\textsf{N}(\mu, \sigma^2) 
\end{aligned}$$
where the  'iid' above the distributed as symbol '$\sim$'  indicates that each of the observations are **i**ndependent of the others (given $\mu$ and $\sigma^2$) and are **i**dentically **d**istributed. 

**Conjugate prior**
Back in \@ref(sec:normal-normal), we found that with normal data, the conjugate prior for $\mu$ when the standard deviation $\sigma$ was known was a normal distribution.  We will build on this  to specify a conditional prior distribution for $\mu$ as 
\begin{equation}
\mu \mid \sigma^2   \sim  \textsf{N}(m_0, \sigma^2/n_0)
(\#eq:04-conjugate-normal)
\end{equation}
with hyper-parameters $m_0$, the prior mean for $\mu$, and $\sigma^2/n_0$ the prior variance.  While previously the variance was a known constant $\tau^2$, replacing $\tau^2$ with a multiple of $\sigma^2$ is needed for representing the joint conjugate prior for the mean and variance.   Because $\sigma$ has the same units as the data, the hyper-parameter $n_0$ is unitless, but is used to express our prior precision about $\mu$ with larger values of $n_0$ indicating more precision and smaller values less precision.  We will see later how the hyper-parameter $n_0$ may be interpreted as a prior sample size.  

As $\sigma^2$ is unknown, a Bayesian would use a
prior distribution to describe the uncertainty about the variance before seeing data.  Since the variance is non-negative, continuous, and with no upper limit,  a gamma distribution is a candidate  prior for the variance, based on the distributions that we have seen so far. However, that choice does not lead to a posterior distribution in the same family or that is recognizable as any common distribution. It turns out that the the inverse of the variance, which is known as the precision, has a conjugate gamma prior distribution.
Letting $\phi = 1/\sigma^2$ denote the precision or inverse variance,  the conjugate prior for $\phi$,
\begin{equation}
\phi \sim \textsf{Gamma}\left(\frac{v_0}{2}, \frac{v_0 s^2_0}{2} \right)
(\#eq:04-conjugate-gamma)
\end{equation}
is a gamma distribution with hyper-parameters $v_0$,  prior degrees of freedom, and $s^2_0$ a prior variance or guess for $\sigma^2$.  Equivalently we may say that the inverse of the variance has a 
$$1/\sigma^2 \sim \textsf{Gamma}(v_0/2, s^2_0 v_0/2)$$

gamma distribution to avoid using a new symbol.  Together the Normal conditional distribution for $\mu$ given $\sigma^2$ in \@ref(eq:04-conjugate-normal)  and the marginal Gamma distribution for $\phi$ in \@ref(eq:04-conjugate-gamma) lead to a joint distribution for the pair $(\mu, \phi)$ that we will call the Normal-Gamma family of distributions:
\begin{equation}(\mu, \phi) \sim \textsf{NormalGamma}(m_0, n_0, s^2_0, v_0)
(\#eq:04-conjugate-normal-gamma)
\end{equation}
with the four hyper-parameters $m_0$, $n_0$, $s^2_0$, and $v_0$.

**Posterior Distribution**

As a conjugate family, the posterior
distribution of the pair of parameters ($\mu, \phi$) is in the same family as the prior distribution when the sample data arise from a  normal distribution, that is the posterior is also Normal-Gamma 
\begin{equation}
(\mu, \phi) \mid \text{data} \sim \textsf{NormalGamma}(m_n, n_n, s^2_n, v_n)
\end{equation}
where the subscript $n$ on the
hyper-parameters indicates the updated values after seeing the $n$ observations. One attraction to conjugate families is there are relatively simple updating rules for obtaining the new hyper-parameters:
\begin{eqnarray*}
m_n & = & \frac{n \bar{Y} + n_0 m_0} {n + n_0}  \\
& \\
n_n & = & n_0 + n  \\
v_n & = & v_0 + n  \\
s^2_n & =  & \frac{1}{v_n}\left[s^2_0 v_0 + s^2 (n-1) + \frac{n_0 n}{n_n} (\bar{Y} - m_0)^2 \right]. 
\end{eqnarray*}
The updated hyper-parameter $m_n$ in the posterior distribution of $\mu$ is the posterior mean, which is a weighted average of the sample mean $\bar{Y}$ and prior mean $m_0$ with weights $n/(n + n_0$ and $n_0/(n + n_0)$ respectively and does not depend on $\sigma^2$.
The posterior sample size $n_n$ is the sum of the prior sample
size $n_n$ and the sample size $n$, representing the combined precision of the estimate for $\mu$.  The posterior degrees of freedom $v_n$ are also increased by adding the  sample size $n$ to the prior degrees of freedom $v_0$. Finally, the posterior variance hyper-parameter $s^2_n$ combines three sources of information about $\sigma$ in terms of sums of squared deviations. **FILL IN MORE DETAILS** The first term in
the square brackets is the sample variance times the sample degrees of
freedom which is the sample sum of squares. The second term represents
the prior sum of squares, while the third term is based on the squared
difference of the sample mean and prior mean.  We then divide by the
posterior degrees of freedom to get the new hyper-parameter.



The joint  Normal-Gamma distribution for the pair $\mu$ and
$\phi$, 
$$(\mu, \phi) \mid \data \sim \NoGa(m_n, n_n, s^2_n, v_n)$$
is equivalent to a **hierarchical model** specified in two stages with $\mu$ given $\sigma$
having a conditional normal distribution 
$$\mu \mid \data, \sigma^2  \sim  \No(m_n, \sigma^2/n_n)$$
and the inverse variance marginally
$$
1/\sigma^2 \mid \data   \sim   \Ga(v_n/2, s^2_n v_n/2) 
$$
having a gamma distribution.  We will see in the next section how this representation is convenient for generating samples from the posterior distribution.


**Marginal Distribution for $\mu$**

We are generally interested in inference about $\mu$ unconditionally
as $\sigma^2$ is unknown. This marginal inference requires the unconditional or marginal distribution of $\mu$ that `averages' over the uncertainty in $\sigma$. For continuous variables like $\sigma$, this averaging is performed by integration leading to the following result:  

$\mu$ given the data is a \index{Student t distribution}
$$ \mu \mid \data \sim \St(v_n, m_n, s^2_n/n_n)  $$ 
with density
\begin{equation}
p(\mu) =\frac{\Gamma\left(\frac{v_n + 1}{2} \right)}
{\sqrt{\pi v_n} \frac{s_n}{\sqrt{n_n}} \,\Gamma\left(\frac{v_n}{2} \right)}
\left(1 + \frac{1}{v_n}\frac{(\mu - m_n)^2} {s^2_n/n_n} \right)^{-\frac{v_n+1}{2}} 
(\#eq:Student-t-density)
\end{equation}
with the degrees of freedom $v_n$, a
location parameter $m_n$ and squared scale parameter that is the
posterior variance parameter divided by the posterior sample size. 
A standard Student $t$ random variable can be obtained by taking  $\mu$ and  subtracting the location $m_n$ and dividing by the scale 
$s_n/\sqrt{n}$:
$$ \frac{\mu - m_n}{s_n/\sqrt{n_n}} \equiv t \sim \St(v_n, 0 , 1)  $$
with degrees of freedom $v_n$, location  $0$ and scale $1$ in the
expression for the density in \@ref(eq:Student-t-density).
This latter representation allows us to use standard statistical functions for posterior inference such as finding credible intervals.

The Student $t$ distribution is similar to the normal distribution as it is symmetric and bell shaped, however, the __tails__ of the distribution are fatter or heavier than the normal distribution. The parameters $m_n$ and $s^2_n$ play similar roles in determining the center and spread of the distribution, as in the Normal distribution, however,  as Student $t$ distributions with degrees of freedom less than 3 do not have a mean or variance, the parameter $m_n$ is called the location or center of the distribution and the $s_n/\sqrt{n}$ is the scale.




**Example**

Let's look at an example based on a sample of total trihalomethanes or TTHM in tap water from a city in NC.  The data can be loaded from the `statsr` package

```{r package, echo=TRUE, warning=FALSE, message=FALSE,  eval=FALSE}
library(statsr)
data(tapwater)
```


```{r loadpackage, warning=FALSE, message=FALSE, echo=FALSE}
suppressWarnings(library(statsr))
data(tapwater)
m_0 = 35; 
n_0 = 25; 
s2_0 = 156.25;
v_0 = n_0 - 1
Y = tapwater$tthm
ybar = mean(Y)
s2 = round(var(Y),1)
n = length(Y)
n_n = n_0 + n
m_n = round((n*ybar + n_0*m_0)/n_n, 1)
v_n = v_0 + n
s2_n = round( ((n-1)*s2 + v_0*s2_0 + n_0*n*(m_0 - ybar)^2/n_n)/v_n, 1)
L = qt(.025, v_n)*sqrt(s2_n/n_n) + m_n
U = qt(.975, v_n)*sqrt(s2_n/n_n) + m_n
```


Using prior information about TTHM from the city, we will use a Normal-Gamma prior distribution,
$\textsf{NormalGamma}(`r m_0`, `r n_0`, `r s2_0`, `r v_0`)$ with
a prior mean of `r m_0` parts per billion, a prior sample
size of `r n_0`, an estimate of the variance of `r s2_0` with degrees of freedom `r v_0`.  In section \@ref(sec:NG-predictive), we will describe how we arrived at these values.

Using the summaries of the data, $\bar{Y} = `r round(ybar, 1)`$,
variance $s^2 = `r s2`$ and sample size of $n = `r n`$ with the
prior hyper-parameters from above, the posterior hyper-parameters are updated as follows:
\begin{eqnarray*}
n_n & = &  `r n_0` +  `r n` = `r n_n`\\
m_n  & = & \frac{`r n` \times`r round(ybar, 1)` + `r n_0` \times`r m_0`}{`r n_n`} = `r round(m_n, 1)`  \\
v_n & = & `r v_0` + `r n` = `r v_n`  \\
s^2_n & = & \frac{(n-1) s^2 + v_0 s^2_0 + n_0 n (m_0 - \bar{Y})^2 /n_n }{v_n}  \\
  & = & \frac{1}{`r v_n`}
     \left[`r n-1` \times `r round(s2, 1)` +
          `r v_0` \times `r s2_0`  +
          \frac{`r n_0` \times `r n`}{`r n_n`} \times (`r m_0` - `r round(ybar, 1)`)^2
\right] = `r round(s2_n, 1)`  \\
\end{eqnarray*}
in the conjugate $\textsf{NormalGamma}(`r m_n`, `r n_n`, `r s2_n`, `r v_n`)$ 
posterior distribution that now summarizes our 
uncertainty about $\mu$ and $\phi$ ($\sigma^2$)  after seeing the data.

We can obtain the updated hyper-parameters in `R` using the following code in `R`
```{r post-NG-TTHM}
# prior hyperparameters
m_0 = 35; n_0 = 25;  s2_0 = 156.25; v_0 = n_0 - 1
# sample summaries
Y = tapwater$tthm
ybar = mean(Y)
s2 = var(Y)
n = length(Y)
# posterior hyperparamters
n_n = n_0 + n
m_n = (n*ybar + n_0*m_0)/n_n
v_n = v_0 + n
s2_n = ((n-1)*s2 + v_0*s2_0 + n_0*n*(m_0 - ybar)^2/n_n)/v_n
```

**Credible intervals for $\mu$**

To find a credible interval for the mean $\mu$, we use the Student $t$
distribution.  Since the distribution of $\mu$ is unimodal and symmetric, the shortest 95 percent credible interval or the **Highest Posterior Density** interval, HPD for short,
```{r tplot, echo=F, warning=FALSE, message=FALSE, fig.width=6, fig.height=6, caption="Highest Posterior Density region for the mean in the total trihalomethanes example."}
require(ggplot2)
out =bayes_inference(tthm, data=tapwater,prior="NG",
                mu_0 = m_0, n_0=n_0, s_0 = sqrt(s2_0), v_0 = v_0,
                method="theoretical", stat="mean", type="ci",show_res=F, show_summ=F, show_plot=F )
 den = out$post_den

 ci = out$ci
 d = data.frame(mu = den$x, dens = den$y)

    d = d[d$dens > 1e-4,]

    li = min(which(d$mu >= ci[1]))
    ui = max(which(d$mu <  ci[2]))

    ci_poly = data.frame(mu = c(d$mu[c(li,li:ui,ui)]),
                         dens = c(0, d$dens[li:ui], 0))


    ci_interval = data.frame(mu = ci, dens = 0)
    pos_plot = ggplot(d, aes_string(x="mu", y="dens")) +
               geom_line() +
               ylab("Density") +
               xlab(expression(mu)) +
               geom_line(data  = ci_interval, size=1.5, colour="orange") +
               geom_point(data = ci_interval, size=2) +
               geom_polygon(data = ci_poly, alpha=0.5) +
               theme(panel.background = element_rect(fill = "transparent", colour = NA),
                     plot.background = element_rect(fill = "transparent", colour = NA)) +
               theme(text = element_text(size=12))
```

```{r tapwater-post-mu, echo=FALSE, fig.width=4,fig.height=4}
print(pos_plot)
```

is the orange interval given by the
Lower endpoint L and upper endpoint U where the probability that mu is
in the interval (L, U) is the shaded area which is equal to zero point
nine five.

using the standardized t distribution and some algebra, these values are
$$
\begin{aligned}
  L & =  m_n + t_{0.025}\sqrt{s^2_n/n_n}    \\
  U & =  m_n + t_{0.975}\sqrt{s^2_n/n_n}
\end{aligned}
$$
or the posterior mean (our point estimate) plus quantiles of the standard $t$ distribution times the scale.  Because of the symmetry in the Student $t$ distribution, the credible interval is $m_n \pm t_{0.975}\sqrt{s^2_n/n_n}$, which should look familiar to expressions for confidence intervals.

Using the following code in `R` the  95\%
credible interval for the tap water data is

```{r CI-tapwater}
m_n + qt(c(0.025, 0.975), v_n)*sqrt(s2_n/n_n)
```




Based on the updated posterior, we find that there is a 95 chance that
the mean TTHM concentration is between `r round(L, 1)`
parts per billion and `r round(U, 1)` parts per billion.

**Summary**
The Normal-Gamma conjugate prior for
inference about an unknown mean and variance for samples from a normal
distribution allows simple expressions for updating prior beliefs given the data.   The joint Normal-Gamma distribution leads to the
Student $t$ distribution for inference about $\mu$ when $\sigma$ is unknown.  The Student $t$ distribution can be used to provide 
credible intervals for $\mu$  using `R` or other software that provides quantiles of a standard $t$ distribution.

For the energetic learner who is comfortable with calculus, the following optional material provides more details on how the posterior distributions were obtained and other results in this section.

For those that are ready to move on, we will introduce Monte Carlo sampling  in the next section; Monte Carlo Sampling is a simulation method that will allow us to approximate distributions of transformations of the parameters without using calculus or change of variables, as well as aid exploratory data analysis of the prior or posterior distribution.



**Details of Results (optional reading)**

TBA

### Monte Carlo Inference  {#sec:NG-MC}

In this section, we will illustrate Monte Carlo sampling from posterior distributions and how it can be used for inference. In Section \@ref(sec:normal-gamma), we can get the posterior distributions for the precision (inverse variance), and the mean given the precision. Then the marginal distribution of the mean can be obtained via integration.

Here is a recap of the joint posterior distribution for the mean $\mu$ and the precision $\phi = 1/\sigma^2$:

* Conditional posterior $\mu \mid \data, \sigma^2  \sim  \No(m_n, \sigma^2/n_n)$
* Marginal posterior $1/\sigma^2 = \phi \mid \data   \sim   \Ga(v_n/2,s^2_n v_n/2)$
* Marginal posterior $\mu \mid \data \sim \St(v_n, m_n, s^2_n/n_n)$

What if we are interested in the distribution of the standard deviation $\sigma$ itself, or other transformations of the parameters? There may not be a closed-form expression for the distributions.

However, it turns out that **Monte Carlo sampling** is an easy way to make inference, when we cannot analytically calculate distributions of parameters, expectations, or probabilities. Monte Carlo methods are computational algorithms that rely on repeated random sampling to calculate numerical results. The name refers to the famous Monte Carlo Casino in Monaco, home to games of chance such as Roulette.

Let's start with a case where we know the posterior distribution.


For posterior inference about the precision $\phi$ using Monte Carlo simulation, we generate $S$ random samples from the posterior distribution: 

$$\phi^{(1)},\phi^{(2)},\cdots,\phi^{(S)} \iid \Ga(v_n/2,s^2_n v_n/2)$$

The term **iid** stands for **i**ndependent and **i**dentically **d**istributed. In other words, the $S$ draws of $\phi$ are independent and identically distributed from the gamma distribution.

Then the empirical distribution of the $S$ samples is used to approximate the actual posterior distribution. From the samples, the sample mean of the draws of $\phi$ can be used to approximate the posterior mean of $\phi$. 

Likewise, we can calculate probabilities, quantiles and other functions using the samples from the posterior distribution. For example, if we want to calculate the posterior expectation of some function of $\phi$, written as $g(\phi)$, we can approximate that by taking the average of the function, and evaluate it at the $S$ draws of $\phi$, written as $\frac{1}{S}\sum^S_{i=1}g(\phi^{(i)})$.

The approximation improves as the size of the Monte Carlo simulation $S$ increases.

$$\frac{1}{S}\sum^S_{i=1}g(\phi^{(i)}) \rightarrow E(g(\phi \mid \data))$$

**Example**

We will apply this with the tap water example. To start, we will set a random seed, which allows the results to be replicated. To generate 1,000 draws from the gamma posterior distribution, we use the `rgamma` function `R` with the posterior hyperparameters from last time.

```{r monte-carlo-example, warning=FALSE, message=FALSE}
set.seed(8675309)
phi = rgamma(1000, shape = v_n/2, rate=s2_n*v_n/2)
```

Figure \@ref(fig:phi-plot) shows the histogram of the 1,000 draws of $\phi$ generated from the Monte Carlo simulation, representing the empirical distribution. The orange line represents the actual gamma posterior density.

```{r phi-plot, warning=FALSE, message=FALSE, echo=FALSE, fig.width=4,fig.height=4, fig.align="center",fig.cap="Empirical distribution of the tap water example"}
library(ggplot2)

d.temp = data.frame(phi)

mcmc.plot = ggplot(d.temp,aes(phi))+ 
      geom_histogram(aes(phi, ..density..),binwidth = 10e-6)+
      geom_density(aes(phi, ..density..),color="orange",size=2)

mcmc.plot
```

Try changing the random seed or increasing the number of simulations, and see how the approximation changes.

We will now use Monte Carlo simulations to approximate the distribution of $\sigma$. Since $\sigma = 1/\sqrt{\phi}$, we can apply the transformation to the 1,000 draws of $\phi$ to obtain a random sample of $\sigma$. We can then estimate the posterior mean of $\sigma$ by calculating the sample mean of the 1,000 draws. 

```{r sigma-post-mean}
sigma = 1/sqrt(phi)
mean(sigma) # posterior mean of sigma
```

Similarly, we can obtain a 95% credible interval for $\sigma$ by finding the sample quantiles of the distribution.

```{r sigma-post-quantile}
quantile(sigma, c(0.025, 0.975))
```

**Summary**

To recap, we have introduced the powerful method of Monte Carlo simulation for posterior inference. Monte Carlo methods provide estimates of expectations, probabilities, and quantiles of distributions from the simulated values. Monte Carlo simulation also allows us to approximate distributions of functions of the parameters, or the transformations of the parameters.

Next we will discuss predictive distributions and show how Monte Carlo simulation may be used to help choose prior hyperparameters, using the prior predictive distribution of data.

### Predictive Distributions {#sec:NG-predictive}

In this section, we will discuss prior and posterior **predictive** distributions of the data and show how Monte Carlo sampling from the prior predictive distribution can help select hyper parameters.

We can obtain the prior predictive distribution of the data, by taking the joint distribution of the data and the parameters in averaging over the possible values of the parameters from the prior.

* Prior:

$$ \begin{align}
\frac{1}{\sigma^2} = \phi &\sim \textsf{Gamma}\left(\frac{v_0}{2}, \frac{v_0 s^2_0}{2} \right) \\
\mu \mid \sigma^2  &\sim  \textsf{N}(m_0, \sigma^2/n_0)
\end{align} $$

* Sampling model:

$$Y_i \mid \mu,\sigma^2 \iid \No(\mu, \sigma^2) $$

* Prior predictive distribution for $Y$:

$$\begin{align}
p(Y) &= \int \int p(Y \mid \mu,\sigma^2) p(\mu \mid \sigma^2) p(\sigma^2) d\mu d\sigma^2 \\
Y &\sim t(v_0, m_0, s_0^2+s_0^2/n_0)
\end{align}$$

This distribution of the observables can be used to help elicit prior hyper parameters as in the tap water example.

A report from the city water department suggests that levels of TTHM are expected to be between 10-60 parts per billion (ppb).

* Set the prior mean $\mu$ to be at the midpoint of the interval: $m_0 = (60+10)/2 = 35$

* Standard deviation: Based on the empirical rule, 95% observations are within $\pm 2\sigma$ of $\mu$, we expect that the range of the data should be $4\sigma$.
 
* Prior estimate of sigma: $s_0 = (60-10)/4 = 12.5$ or $s_0^2 = [(60-10)/4]^2 = 156.25$

To complete the specification, we also need to choose the prior sample size $n_0$ and degrees of freedom $v_0$. As the degrees of freedom of the variance are $n-1$, we set $v_0 = n_0 - 1$. We will draw samples from the prior predictive distribution and modify $n_0$ so that the simulated data agree with our prior assumptions.

The following `R` code shows a simulation from the predictive distribution with the prior sample size of 2. Please note that the number of Monte Carlo simulations should not be confused with the prior sample size $n_0$.

We begin by simulating $\phi$, transfering $\phi$ to calculate $\sigma$, and then simulating values of $\mu$. Finally, the simulated values of $\mu,\sigma$ are used to generate possible values of TTHM denoted by $Y$. 

```{r predictive-TTHM}
m_0 = (60+10)/2; s2_0 = ((60-10)/4)^2;
n_0 = 2; v_0 = n_0 - 1
set.seed(1234)
phi = rgamma(1000, v_0/2, s2_0*v_0/2)
sigma = 1/sqrt(phi)
mu = rnorm(1000, mean=m_0, sd=sigma/(sqrt(n_0)))
y = rnorm(1000, mu, sigma)
quantile(y, c(0.025,0.975))
```

This forward simulation propagates uncertainty in $\mu,\sigma$ to the prior predictive distribution of the data. Calculating the sample quantiles from the samples of the prior predictive for $Y$, we see that the 95% predictive interval includes negative values. Since TTHM is non-negative, we need to adjust $n_0$ and repeat.

### Reference Priors  {#sec:NG-reference}

### Mixtures of Conjugate Priors  {#sec:NG-Cauchy}

### MCMC  {#sec:NG-MCMC}


 
